---
title: "Mouse Androgen Clock Construction"
author: "Victoria Sugrue"
date: 'Last updated: 30 May 2023'
output:
  word_document:
    toc: no
    toc_depth: '6'
  html_document:
    toc: no
    toc_float: no
    toc_depth: 6
  pdf_document:
    toc: no
    toc_depth: 5
    df_print: kable
    latex_engine: xelatex
subtitle: ' '
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r}
library(ggplot2)
library(caret)
library(dplyr)
library(Manu)
library(knitr)
library(sjPlot)
library(lmtest)
library(Amelia)
library(mlbench)
library(ggfortify)
```

# Data preparation
```{r}
mouse_data <- read.csv("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Bioinformatics/Data/Mouse_ArrayData.csv")

# Setting AndroAge with predicted time of puberty onset = age minus 6 or 8 weeks years
mouse_data$YPMP_6 <- mouse_data$Age-0.115068
mouse_data$YPMP_6[mouse_data$Sex=="F"] <- 0
mouse_data$YPMP_8 <- mouse_data$Age-0.153425
mouse_data$YPMP_8[mouse_data$Sex=="F"] <- 0

# Adjust to 0 where negative
mouse_data$YPMP_6[mouse_data$YPMP_6 < 0] <- 0
mouse_data$YPMP_8[mouse_data$YPMP_8 < 0] <- 0

# Change sex labels
mouse_data$Sex[mouse_data$Sex=="F"] <- "Female"
mouse_data$Sex[mouse_data$Sex=="M"] <- "Male"


# Subset by tissue
unique(mouse_data$Tissue)
mouse_cortex <- mouse_data %>% filter(Tissue=="Cortex")
mouse_striatum <- mouse_data %>% filter(Tissue=="Striatum")
mouse_blood <- mouse_data %>% filter(Tissue=="Blood")
mouse_cerebellum <- mouse_data %>% filter(Tissue=="Cerebellum")
mouse_liver <- mouse_data %>% filter(Tissue=="Liver")
mouse_muscle <- mouse_data %>% filter(Tissue=="Muscle")
mouse_kidney <- mouse_data %>% filter(Tissue=="Kidney")
mouse_hippocampus <- mouse_data %>% filter(Tissue=="Hippocampus")
mouse_tail <- mouse_data %>% filter(Tissue=="Tail")
```

# First, lets see if methylation levels and sex (taking castration into account) are correlated.

# Cortex

```{r}
# Convert the "sex" variable to a numeric variable
mouse_cortex$Sex_numeric <- ifelse(mouse_cortex$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_cortex <- cor.test(mouse_cortex$cg21524116, mouse_cortex$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_cortex)
```

# Striatum

```{r}
# Convert the "sex" variable to a numeric variable
mouse_striatum$Sex_numeric <- ifelse(mouse_striatum$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_striatum <- cor.test(mouse_striatum$cg21524116, mouse_striatum$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_striatum)
```

# Blood

```{r}
# Convert the "sex" variable to a numeric variable
mouse_blood$Sex_numeric <- ifelse(mouse_blood$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_blood <- cor.test(mouse_blood$cg21524116, mouse_blood$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_blood)
```

# Cerebellum

```{r}
# Convert the "sex" variable to a numeric variable
mouse_cerebellum$Sex_numeric <- ifelse(mouse_cerebellum$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_cerebellum <- cor.test(mouse_cerebellum$cg21524116, mouse_cerebellum$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_cerebellum)
```

# Liver

```{r}
# Convert the "sex" variable to a numeric variable
mouse_liver$Sex_numeric <- ifelse(mouse_liver$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_liver <- cor.test(mouse_liver$cg21524116, mouse_liver$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_liver)
```

# Muscle

```{r}
# Convert the "sex" variable to a numeric variable
mouse_muscle$Sex_numeric <- ifelse(mouse_muscle$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_muscle <- cor.test(mouse_muscle$cg21524116, mouse_muscle$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_muscle)
```

# Kidney

```{r}
# Convert the "sex" variable to a numeric variable
mouse_kidney$Sex_numeric <- ifelse(mouse_kidney$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_kidney <- cor.test(mouse_kidney$cg21524116, mouse_kidney$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_kidney)
```

# Hippocampus

```{r}
# Convert the "sex" variable to a numeric variable
mouse_hippocampus$Sex_numeric <- ifelse(mouse_hippocampus$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_hippocampus <- cor.test(mouse_hippocampus$cg21524116, mouse_hippocampus$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_hippocampus)
```

# Tail

```{r}
# Convert the "sex" variable to a numeric variable
mouse_tail$Sex_numeric <- ifelse(mouse_tail$Sex == "Male", 0, 1)

# Use cor.test to test for correlation between DNA methylation and sex
cor_tail <- cor.test(mouse_tail$cg21524116, mouse_tail$Sex_numeric, method="spearman")

# Print the results of the correlation test
print(cor_tail)
```

# A birds-eye view of the data 

```{r}
# Age distribution
age_dist <- ggplot(mouse_data, aes(x = Age)) +
  geom_histogram(aes(fill = Tissue), position = "dodge", alpha = 1, bins = 15) +
  #scale_fill_manual(values = get_pal("Takahe")) +
  theme_light() +
  facet_grid(~Sex) +
  labs(title = "Distribution of age by sex (Mouse, array)", fill = "Sex", x="Age (Years)") +
  ylab("Frequency")
age_dist

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/Mouse/age_distribution_clockMouse.png",age_dist, width=10, height = 8, dpi=1000)
```


```{r}
#or, by tissue
age_dist_tis <- ggplot(mouse_data, aes(x = Age)) +
  geom_histogram(aes(fill = Sex), position = "dodge", alpha = 1, bins = 15) +
  scale_fill_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  facet_wrap(~Tissue, ncol=3, scales="free") +
  labs(title = "Distribution of age by sex (Mouse, array)", fill = "Sex", x="Age (Years)") +
  ylab("Frequency")
age_dist_tis

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/Mouse/age_distribution_clockMouse_tissue.png",age_dist_tis, width=10, height = 7, dpi=1000)

```


# Plotting methylation data

```{r}
MKLN1_mice <- ggplot(mouse_data, aes(x=Age, y=cg21524116*100, colour=Sex)) +
  geom_jitter(width=0.05, size=2.5, alpha=2/3) +
  geom_smooth(method="lm") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  facet_wrap(~Tissue, ncol=3, scales="free") +
  labs(colour="Sex", y="Methylation (%)", x="Age (Years)")
MKLN1_mice

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/Mouse/MKLN1_meth_mouse.png",MKLN1_mice, width=10, height = 7, dpi=1000)
```

4 options of clocks to test: 
Mouse Tail; onset of puberty 6 weeks
Mouse Tail, onset of puberty 8 weeks
Mouse Muscle, onset of puberty 6 weeks
Mouse Muscle, onset of puberty 8 weeks

##### Tail 6 weeks #####
# Fit androgen clock to mouse tail, using 6 weeks as puberty onset
```{r}
male_tail <- mouse_tail %>% filter(Sex=="Male")

#onset of puberty set at 6 weeks
tail6 <- lm(YPMP_6 ~ cg21524116, data=male_tail)
summary(tail6)
```

# Diagnostic plot

```{r}
pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/mouse_tail6_ggplot_diagnostics.pdf", width=10, height=8)

autoplot(tail6, which=c(1,2,3,5), size=3, alpha=4/5, colour="#313657", smooth.colour="#DD3C51", smooth.linetype="dashed", label.size = 3) + theme_light() + theme(plot.title = element_text(color = "white"))

dev.off()
```

```{r}
fitted <- as.data.frame(tail6$fitted.values)
fitted["YPMP_6"] <- male_tail$YPMP_6
fitted$YPMP_6 <- as.numeric(fitted$YPMP_6)
fitted$Fit_YPMP_6 <- as.numeric(fitted$`tail6$fitted.values`)
```

# Calculating error
## All

```{r}
MAE_tail6 <- as.data.frame(Metrics::mae(actual=fitted$YPMP_6, predicted=fitted$`tail6$fitted.values`))
MAE_tail6 # 0.250
```

```{r}
cor(fitted$YPMP_6, fitted$Fit_YPMP_6, method="pearson")
```



# Testing the model fit further: 

# Test density/normality of residuals - looking for normally distributed. 
```{r}
d<-density(tail6[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

```{r}
ggplot(data = male_tail, aes(x = tail6$residuals)) +
    geom_histogram(fill = 'grey40', color = 'black', bins=20) +
    labs(title = 'Histogram of Residuals, Tail, 6 weeks', x = 'Residuals', y = 'Frequency')
```
# Breusch-Pagan test of variance (homo/heteroscedasticity)

```{r}
plot(tail6$residuals)
bptest(tail6)
```

# p value >0.05 suggests that we have homoscedasticity. good! 

# Shapiro-Wilk test of normality

```{r}
shapiro.test(tail6$residuals)
```

# p value <0.05 suggests data isn't normally distributed. not good!

# Leave-one-out cross validation

```{r}
male_tail$cg21524116 <- as.numeric(male_tail$cg21524116)
male_tail$YPMP_6 <- as.numeric(male_tail$YPMP_6)
ctrl <- trainControl(method="LOOCV")
tail6LOO <- train(YPMP_6 ~ cg21524116, data=male_tail, method="lm", trControl=ctrl)
summary(tail6LOO)
```

```{r}
print(tail6LOO)
```

```{r}
fittedLOO <- as.data.frame(tail6LOO$pred)
fittedLOO["YPMP_6"] <- male_tail$YPMP_6

tail6LOO_MAE <- as.data.frame(Metrics::mae(actual=fittedLOO$YPMP_6, predicted=fittedLOO$pred))
tail6LOO_MAE # 0.280

cor(fittedLOO$obs, fittedLOO$pred, method="pearson")

tail6fit <- ggplot(fitted, aes(x=YPMP_6, y=Fit_YPMP_6)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
    labs(title = "Tail, 6 weeks",
              subtitle = "MAE=0.280, cor=0.770") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
tail6fit
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/tail6fit.png",tail6fit, width=8, height = 7, dpi=1000)
```


# Now, to include females.
# First, apply the formula of the male androgen clock to this group. What happens? 

```{r}
female_tail <- mouse_tail %>% filter(Sex=="Female")
female_tail$Model_AndroAge <- predict(tail6, newdata = female_tail)

female_MAE_tail6 <- as.data.frame(Metrics::mae(actual=female_tail$YPMP_6, predicted=female_tail$Model_AndroAge))
female_MAE_tail6
# Female MAE = 0.12
```
















##### Tail 8 weeks #####

# Fit androgen clock to mouse tail, using 8 weeks as puberty onset
```{r}
male_tail <- mouse_tail %>% filter(Sex=="Male")

#onset of puberty set at 6 weeks
tail8 <- lm(YPMP_8 ~ cg21524116, data=male_tail) 
summary(tail8)

```

# Diagnostic plots

```{r}
pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/mouse_tail8_ggplot_diagnostics.pdf", width=10, height=8)

autoplot(tail8, which=c(1,2,3,5), size=3, alpha=4/5, colour="#313657", smooth.colour="#DD3C51", smooth.linetype="dashed", label.size = 3) + theme_light() + theme(plot.title = element_text(color = "white"))

dev.off()
```

```{r}
fitted <- as.data.frame(tail8$fitted.values)
fitted["YPMP_8"] <- male_tail$YPMP_8
fitted$YPMP_8 <- as.numeric(fitted$YPMP_8)
fitted$Fit_YPMP_8 <- as.numeric(fitted$`tail8$fitted.values`)
```

# Calculating error
## All

```{r}
MAE_tail8 <- as.data.frame(Metrics::mae(actual=fitted$YPMP_8, predicted=fitted$`tail8$fitted.values`))
MAE_tail8 # 0.253
```
```{r}
cor(fitted$YPMP_8, fitted$Fit_YPMP_8, method="pearson")
```

# Testing the model fit further: 

# Test density/normality of residuals - looking for normally distributed. 
```{r}
d<-density(tail8[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

```{r}
ggplot(data = male_tail, aes(x = tail8$residuals)) +
    geom_histogram(fill = 'grey40', color = 'black', bins=20) +
    labs(title = 'Histogram of Residuals, Tail, 8 weeks', x = 'Residuals', y = 'Frequency')
```
# Breusch-Pagan test of variance (homo/heteroscedasticity)

```{r}
plot(tail8$residuals)
bptest(tail8)
```

# p value >0.05 suggests that we have homoscedasticity. good! 

# Shapiro-Wilk test of normality

```{r}
shapiro.test(tail8$residuals)
```

# p value <0.05 suggests data is normally distributed.  good! (just)

# Leave-one-out cross validation

```{r}
male_tail$cg21524116 <- as.numeric(male_tail$cg21524116)
male_tail$YPMP_8 <- as.numeric(male_tail$YPMP_8)
ctrl <- trainControl(method="LOOCV")
tail8LOO <- train(YPMP_8 ~ cg21524116, data=male_tail, method="lm", trControl=ctrl)
summary(tail8LOO)
```

```{r}
print(tail8LOO)
```

```{r}
fittedLOO <- as.data.frame(tail8LOO$pred)
fittedLOO["YPMP_8"] <- male_tail$YPMP_8

tail8LOO_MAE <- as.data.frame(Metrics::mae(actual=fittedLOO$YPMP_8, predicted=fittedLOO$pred))
tail8LOO_MAE # 0.283

cor(fittedLOO$pred, fittedLOO$obs, method="pearson")


tail8fit <- ggplot(fitted, aes(x=YPMP_8, y=Fit_YPMP_8)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
    labs(title = "Tail, 8 weeks",
              subtitle = "MAE=0.283, cor=0.760") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
tail8fit
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/tail8fit.png",tail8fit, width=8, height = 7, dpi=1000)
```


# Now, to include females.
# First, apply the formula of the male androgen clock to this group. What happens? 

```{r}
female_tail <- mouse_tail %>% filter(Sex=="Female")
female_tail$Model_AndroAge <- predict(tail8, newdata = female_tail)

female_MAE_tail8 <- as.data.frame(Metrics::mae(actual=female_tail$YPMP_8, predicted=female_tail$Model_AndroAge))
female_MAE_tail8
# Female MAE = 0.11
```

# Fit androgen clock to mouse muscle, using 6 weeks as puberty onset
```{r}
male_muscle <- mouse_muscle %>% filter(Sex=="Male")

#onset of puberty set at 6 weeks
muscle6 <- lm(YPMP_6 ~ cg21524116, data=male_muscle) 
summary(muscle6)
```

# Diagnostic plots

```{r}
pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/mouse_muscle6_ggplot_diagnostics.pdf", width=10, height=8)

autoplot(muscle6, which=c(1,2,3,5), size=3, alpha=4/5, colour="#313657", smooth.colour="#DD3C51", smooth.linetype="dashed", label.size = 3) + theme_light() + theme(plot.title = element_text(color = "white"))

dev.off()
```


```{r}
fitted <- as.data.frame(muscle6$fitted.values)
fitted["YPMP_6"] <- male_muscle$YPMP_6
fitted$YPMP_6 <- as.numeric(fitted$YPMP_6)
fitted$Fit_YPMP_6 <- as.numeric(fitted$`muscle6$fitted.values`)
```

# Calculating error
## All

```{r}
MAE_muscle6 <- as.data.frame(Metrics::mae(actual=fitted$YPMP_6, predicted=fitted$`muscle6$fitted.values`))
MAE_muscle6 # 0.516
```
```{r}
cor(fitted$YPMP_6, fitted$Fit_YPMP_6, method="pearson")

```



# Testing the model fit further: 

# Test density/normality of residuals - looking for normally distributed. 
```{r}
d<-density(muscle6[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

```{r}
ggplot(data = male_muscle, aes(x = muscle6$residuals)) +
    geom_histogram(fill = 'grey40', color = 'black', bins=20) +
    labs(title = 'Histogram of Residuals, Muscle, 6 weeks', x = 'Residuals', y = 'Frequency')
```
# Breusch-Pagan test of variance (homo/heteroscedasticity)

```{r}
plot(muscle6$residuals)
bptest(muscle6)
```

# p value >0.05 suggests that we have homoscedasticity. good! 

# Shapiro-Wilk test of normality

```{r}
shapiro.test(muscle6$residuals)
```

# p value <0.05 suggests data isn't normally distributed. not good!

# Leave-one-out cross validation

```{r}
male_muscle$cg21524116 <- as.numeric(male_muscle$cg21524116)
male_muscle$YPMP_6 <- as.numeric(male_muscle$YPMP_6)
ctrl <- trainControl(method="LOOCV")
muscle6LOO <- train(YPMP_6 ~ cg21524116, data=male_muscle, method="lm", trControl=ctrl)
summary(muscle6LOO)
```

```{r}
print(muscle6LOO)
```

```{r}
fittedLOO <- as.data.frame(muscle6LOO$pred)
fittedLOO["YPMP_6"] <- male_muscle$YPMP_6

muscle6LOO_MAE <- as.data.frame(Metrics::mae(actual=fittedLOO$YPMP_6, predicted=fittedLOO$pred))
muscle6LOO_MAE # 0.55

cor(fittedLOO$pred, fittedLOO$obs, method="pearson")


muscle6fit <- ggplot(fitted, aes(x=YPMP_6, y=Fit_YPMP_6)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
    labs(title = "Muscle, 6 weeks",
              subtitle = "MAE=0.555, cor=0.394") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
muscle6fit
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/muscle6fit.png",muscle6fit, width=8, height = 7, dpi=1000)
```


# Now, to include females.
# First, apply the formula of the male androgen clock to this group. What happens? 

```{r}
female_muscle <- mouse_muscle %>% filter(Sex=="Female")
female_muscle$Model_AndroAge <- predict(muscle6, newdata = female_muscle)

female_MAE_muscle6 <- as.data.frame(Metrics::mae(actual=female_muscle$YPMP_6, predicted=female_muscle$Model_AndroAge))
female_MAE_muscle6
# Female MAE = 0.24
```



























# Fit androgen clock to mouse muscle, using 8 weeks as puberty onset
```{r}
#onset of puberty set at 8 weeks
muscle8 <- lm(YPMP_8 ~ cg21524116, data=male_muscle) 
summary(muscle8)
```

# Diagnostic plots

```{r}
pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/mouse_muscle8_ggplot_diagnostics.pdf", width=10, height=8)

autoplot(muscle8, which=c(1,2,3,5), size=3, alpha=4/5, colour="#313657", smooth.colour="#DD3C51", smooth.linetype="dashed", label.size = 3) + theme_light() + theme(plot.title = element_text(color = "white"))

dev.off()
```

```{r}
fitted <- as.data.frame(muscle8$fitted.values)
fitted["YPMP_8"] <- male_muscle$YPMP_8
fitted$YPMP_8 <- as.numeric(fitted$YPMP_8)
fitted$Fit_YPMP_8 <- as.numeric(fitted$`muscle8$fitted.values`)
```

# Calculating error
## All

```{r}
MAE_muscle8 <- as.data.frame(Metrics::mae(actual=fitted$YPMP_8, predicted=fitted$`muscle8$fitted.values`))
MAE_muscle8 # 0.52
```
```{r}
cor(fitted$YPMP_8, fitted$Fit_YPMP_8, method="pearson")

```

# Testing the model fit further: 

# Test density/normality of residuals - looking for normally distributed. 
```{r}
d<-density(muscle8[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

```{r}
ggplot(data = male_muscle, aes(x = muscle8$residuals)) +
    geom_histogram(fill = 'grey40', color = 'black', bins=20) +
    labs(title = 'Histogram of Residuals, Muscle, 8 weeks', x = 'Residuals', y = 'Frequency')
```
# Breusch-Pagan test of variance (homo/heteroscedasticity)

```{r}
plot(muscle8$residuals)
bptest(muscle8)
```

# p value >0.05 suggests that we have homoscedasticity. good! 

# Shapiro-Wilk test of normality

```{r}
shapiro.test(muscle8$residuals)
```

# p value <0.05 suggests data isn't normally distributed. not good!

# Leave-one-out cross validation

```{r}
male_muscle$cg21524116 <- as.numeric(male_muscle$cg21524116)
male_muscle$YPMP_8 <- as.numeric(male_muscle$YPMP_8)
ctrl <- trainControl(method="LOOCV")
muscle8LOO <- train(YPMP_8 ~ cg21524116, data=male_muscle, method="lm", trControl=ctrl)
summary(muscle8LOO)
```

```{r}
print(muscle8LOO)
```

```{r}
fittedLOO <- as.data.frame(muscle8LOO$pred)
fittedLOO["YPMP_8"] <- male_muscle$YPMP_8

muscle8LOO_MAE <- as.data.frame(Metrics::mae(actual=fittedLOO$YPMP_8, predicted=fittedLOO$pred))
muscle8LOO_MAE # 0.55

cor(fittedLOO$obs, fittedLOO$pred, method="pearson")


muscle8fit <- ggplot(fitted, aes(x=YPMP_8, y=Fit_YPMP_8)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
    labs(title = "Muscle, 8 weeks",
              subtitle = "MAE=0.555, cor=0.381") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
muscle8fit
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/muscle8fit.png",muscle8fit, width=8, height = 7, dpi=1000)
```


# Now, to include females.
# First, apply the formula of the male androgen clock to this group. What happens? 

```{r}
female_muscle <- mouse_muscle %>% filter(Sex=="Female")
female_muscle$Model_AndroAge <- predict(muscle8, newdata = female_muscle)

female_MAE_muscle8 <- as.data.frame(Metrics::mae(actual=female_muscle$YPMP_8, predicted=female_muscle$Model_AndroAge))
female_MAE_muscle8
# Female MAE = 0.24
```



## Using the tail clock, 8 weeks puberty.

```{r}
mouse_tail$Model_AndroAge <- predict(tail8, newdata = mouse_tail)

MAE_tail8_all <- as.data.frame(Metrics::mae(actual=mouse_tail$YPMP_6, predicted=mouse_tail$Model_AndroAge))
MAE_tail8_all
```


```{r}
cor(mouse_tail$YPMP_6, mouse_tail$Model_AndroAge, method="pearson")


obs_exp_mouseTail8 <- ggplot(mouse_tail, aes(x=YPMP_6, y=Model_AndroAge, colour=Sex)) +
 geom_point(size=4, alpha=3/4) + 
  geom_smooth(method="lm", se=F, colour="#1F6683") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2)]) +
  theme_light() +
    geom_abline(linetype="dashed") +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
    labs(title = "Mouse androgen clock fit (Tail, 8 weeks), trained on all males",
              subtitle = "MAE=0.182, cor=0.861") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
obs_exp_mouseTail8


ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/Obs_Exp_AndroAge_mouseTail8.png",obs_exp_mouseTail8, width=7, height = 6, dpi=1000)

```


```{r}
write.csv(mouse_tail, "/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Bioinformatics/Data/mouse_data_modelled_tail8.csv")
```


```{r}
mousefit <- cowplot::plot_grid(tail6fit, tail8fit, muscle6fit, muscle8fit, nrow=2, labels=c("A", "B", "C", "D"))
mousefit

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/MouseModelFit_Tissue_AndroAge.png",mousefit, width=8, height = 7, dpi=1000)

```


# Applying the model to the DHT mouse data
```{r}
DHTdata <- read.csv("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Bioinformatics/Data/DHTmice.csv")
DHTdata$cg21524116 <- DHTdata$Methylation/100

DHTtail <- DHTdata %>% filter(Tissue=="Tail")

DHTtail$Model_AndroAge <- predict(tail8, newdata = DHTtail)

MAE_DHTtail_all <- as.data.frame(Metrics::mae(actual=0, predicted=DHTtail$Model_AndroAge))
MAE_DHTtail_all
# 1.206232
```
## controls only
```{r}
DHTtail_control <- DHTtail %>% filter(Treatment=="Blank")

DHTtail_control$Model_AndroAge <- predict(tail8, newdata = DHTtail_control)

MAE_DHTtail_control <- as.data.frame(Metrics::mae(actual=0, predicted=DHTtail_control$Model_AndroAge))
MAE_DHTtail_control
# 0.488161
```

## DHT only
```{r}
DHTtail_DHT <- DHTtail %>% filter(Treatment=="DHT")

DHTtail_DHT$Model_AndroAge <- predict(tail8, newdata = DHTtail_DHT)

MAE_DHTtail_DHT <- as.data.frame(Metrics::mae(actual=0, predicted=DHTtail_DHT$Model_AndroAge))
MAE_DHTtail_DHT
# 2.026885
```
















