---
title: "Sheep Androgen Clock Construction - Cross Platform & Amplicon/Array Comparisons"
author: "Victoria Sugrue"
date: 'Last updated: 16 May 2023'
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 6
  pdf_document:
    toc: yes
    toc_depth: 5
    df_print: kable
    latex_engine: xelatex
subtitle: ' '
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r}
library(ggplot2)
library(caret)
library(dplyr)
library(Manu)
library(knitr)
library(sjPlot)
library(lmtest)
library(Amelia)
library(mlbench)
library(ggfortify)
library(ggpubr)
```

# Data preparation, amplicon
```{r}
sheep_data_amp <- read.csv("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Bioinformatics/Data/SheepSampleSheet.csv")

# Checking for missingness
#pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/sheepdata_missingness.pdf") 
missmap(sheep_data_amp, col=c("blue", "red"), legend=T)
#dev.off()

# Setting AndroAge with predicted time of puberty onset = age minus 0.5 years (i.e estimate to reach puberty at 6 months)
sheep_data_amp$YPMP <- sheep_data_amp$Age-0.5
sheep_data_amp$YPMP[sheep_data_amp$Castration_Sex=="Wether"] <- 0
sheep_data_amp$YPMP[sheep_data_amp$Castration_Sex=="Ewe"] <- 0

# Creating a new column for methylation in the form of a beta value
sheep_data_amp$Beta <- sheep_data_amp$CG_Methylation/100

# Subset by sex/castration status
ram_data_amp <- sheep_data_amp %>% filter(Castration_Sex=="Ram")
wether_data_amp <- sheep_data_amp %>% filter(Castration_Sex=="Wether")
ewe_data_amp <- sheep_data_amp %>% filter(Castration_Sex=="Ewe")

# Subset based on error allowance
sheep_data_amp_5 <- sheep_data_amp %>% filter(CG_Calls>=497)
sheep_data_amp_25 <- sheep_data_amp %>% filter(CG_Calls>=1990)
sheep_data_amp_2 <- sheep_data_amp %>% filter(CG_Calls>=3110)
sheep_data_amp_15 <- sheep_data_amp %>% filter(CG_Calls>=5529)
sheep_data_amp_1 <- sheep_data_amp %>% filter(CG_Calls>=12440)

```

# Fitting the Androgen Clock to rams

```{r}
ram_data_amp_5 <- ram_data_amp %>% filter(CG_Calls>=497)
ram_data_amp_25 <- ram_data_amp %>% filter(CG_Calls>=1990)
ram_data_amp_15 <- ram_data_amp %>% filter(CG_Calls>=5529)
ram_data_amp_2 <- ram_data_amp %>% filter(CG_Calls>=3110)
ram_data_amp_1 <- ram_data_amp %>% filter(CG_Calls>=12440)
```

# Running model, 5% error

```{r}
# Run model, 5% error
model_5p <- lm(YPMP ~ Beta, data=ram_data_amp_5)
summary(model_5p)
```

# Running model, 2.5% error

```{r}
# Run model, 2.5% error
model_25p <- lm(YPMP ~ Beta, data=ram_data_amp_25)
summary(model_25p)
```

# Data preparation, array
```{r}
#read in minfi normalised beta values
sheep_array_data2 <- read.csv("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Bioinformatics/Data/top10asDMP.csv", header=FALSE)

#transpose the dataframe and remove unneeded columns
sheep_array_data <- as.data.frame(t(sheep_array_data2[,-c(1:2)]))

#read in file for headers and transpose it, assign the headers to the minfi normalised data
headers <- read.csv("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Bioinformatics/Data/top10asDMP_headers.csv", header=FALSE)
theaders <- as.list(t(headers))
names(sheep_array_data) <- theaders

sheep_array_data_ear <- sheep_array_data %>% filter(!(Sex=="Blood"))



# Checking for missingness
#pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/sheepdata_missingness.pdf") 
missmap(sheep_array_data, col=c("blue", "red"), legend=T)
missmap(sheep_array_data_ear, col=c("blue", "red"), legend=T)

#dev.off()

# Setting AndroAge with predicted time of puberty onset = age minus 0.5 years (i.e estimate to reach puberty at 6 months)
sheep_array_data_ear$YPMP <- as.numeric(sheep_array_data_ear$Age)-0.5
sheep_array_data_ear$YPMP[sheep_array_data_ear$Sex=="Wether"] <- 0
sheep_array_data_ear$YPMP[sheep_array_data_ear$Sex=="Ewe"] <- 0

# Subset by sex/castration status
ram_array_ear_data <- sheep_array_data_ear %>% filter(Sex=="Ram")
wether_array_ear_data <- sheep_array_data_ear %>% filter(Sex=="Wether")
ewe_array_ear_data <- sheep_array_data_ear %>% filter(Sex=="Ewe")

```

# Fitting model to array data, sheep ear only

```{r}
# Run model, array data, training on rams
model_array1 <- lm(YPMP ~ as.numeric(cg21524116), data=ram_array_ear_data)
summary(model_array1)
```







ARRAY
######## applying amplicon data to array clock!
#### 5 % error amplicon data
```{r}
sheep_data_amp_5$cg21524116 <- sheep_data_amp_5$Beta

ram_data_amp_5 <- sheep_data_amp_5 %>% filter(Castration_Sex=="Ram")
ewe_data_amp_5 <- sheep_data_amp_5 %>% filter(Castration_Sex=="Ewe")
wether_data_amp_5 <- sheep_data_amp_5 %>% filter(Castration_Sex=="Wether")


sheep_data_amp_5$Array_Model_AndroAge <- predict(model_array1, newdata = sheep_data_amp_5)

sheep_amplicon_MAE_model_array1 <- as.data.frame(Metrics::mae(actual=sheep_data_amp_5$YPMP, predicted=sheep_data_amp_5$Array_Model_AndroAge))
sheep_amplicon_MAE_model_array1
# OVERALL MAE = 0.723 years

sheep_data_amp_5 <- sheep_data_amp_5 %>% 
  arrange(Age) %>% 
  mutate(index = row_number())

cor(sheep_data_amp_5$YPMP, sheep_data_amp_5$Array_Model_AndroAge, method="pearson")
# 0.931

compare_array_5 <- ggplot(sheep_data_amp_5, aes(x=YPMP, y=Array_Model_AndroAge, colour=Castration_Sex)) + 
  geom_point(size=4, alpha=3/4) +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  geom_smooth(method="lm", se=F, colour="#1F6683") +
  geom_abline(linetype="dashed") +
  theme_light() +
    ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Train: array, Test: ±5% error amplicon",
              subtitle = "MAE=0.723, cor=0.931") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
compare_array_5
```


```{r}
ram_data_amp_5$Array_Model_AndroAge <- predict(model_array1, newdata = ram_data_amp_5)

ram_amplicon_MAE_model_array1 <- as.data.frame(Metrics::mae(actual=ram_data_amp_5$YPMP, predicted=ram_data_amp_5$Array_Model_AndroAge))
ram_amplicon_MAE_model_array1
# Ram MAE = 0.807 years

ewe_data_amp_5$Array_Model_AndroAge <- predict(model_array1, newdata = ewe_data_amp_5)

ewe_amplicon_MAE_model_array1 <- as.data.frame(Metrics::mae(actual=ewe_data_amp_5$YPMP, predicted=ewe_data_amp_5$Array_Model_AndroAge))
ewe_amplicon_MAE_model_array1
# Ewe MAE = 0.736 years

wether_data_amp_5$Array_Model_AndroAge <- predict(model_array1, newdata = wether_data_amp_5)

wether_amplicon_MAE_model_array1 <- as.data.frame(Metrics::mae(actual=wether_data_amp_5$YPMP, predicted=wether_data_amp_5$Array_Model_AndroAge))
wether_amplicon_MAE_model_array1
# Wether MAE = 0.560 years
```



#### 2.5 % error amplicon data
```{r}
sheep_data_amp_25$cg21524116 <- sheep_data_amp_25$Beta

ram_data_amp_25 <- sheep_data_amp_25 %>% filter(Castration_Sex=="Ram")
ewe_data_amp_25 <- sheep_data_amp_25 %>% filter(Castration_Sex=="Ewe")
wether_data_amp_25 <- sheep_data_amp_25 %>% filter(Castration_Sex=="Wether")


sheep_data_amp_25$Array_Model_AndroAge <- predict(model_array1, newdata = sheep_data_amp_25)

sheep_amplicon_MAE_model_array1 <- as.data.frame(Metrics::mae(actual=sheep_data_amp_25$YPMP, predicted=sheep_data_amp_25$Array_Model_AndroAge))
sheep_amplicon_MAE_model_array1
# OVERALL MAE = 0.708 years

sheep_data_amp_25 <- sheep_data_amp_25 %>% 
  arrange(Age) %>% 
  mutate(index = row_number())

cor(sheep_data_amp_25$YPMP, sheep_data_amp_25$Array_Model_AndroAge, method="pearson")
# 0.928

compare_array_25 <- ggplot(sheep_data_amp_25, aes(x=YPMP, y=Array_Model_AndroAge, colour=Castration_Sex)) + 
  geom_point(size=4, alpha=3/4) +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  geom_smooth(method="lm", se=F, colour="#1F6683") +
  geom_abline(linetype="dashed") +
  theme_light() +
    ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Train: array, Test: ±2.5% error amplicon",
              subtitle = "MAE=0.708, cor=0.928") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
compare_array_25
```


```{r}
ram_data_amp_25$Array_Model_AndroAge <- predict(model_array1, newdata = ram_data_amp_25)

ram_amplicon_MAE_model_array1 <- as.data.frame(Metrics::mae(actual=ram_data_amp_25$YPMP, predicted=ram_data_amp_25$Array_Model_AndroAge))
ram_amplicon_MAE_model_array1
# Ram MAE = 0.735 years

ewe_data_amp_25$Array_Model_AndroAge <- predict(model_array1, newdata = ewe_data_amp_25)

ewe_amplicon_MAE_model_array1 <- as.data.frame(Metrics::mae(actual=ewe_data_amp_25$YPMP, predicted=ewe_data_amp_25$Array_Model_AndroAge))
ewe_amplicon_MAE_model_array1
# Ewe MAE = 0.787 years

wether_data_amp_25$Array_Model_AndroAge <- predict(model_array1, newdata = wether_data_amp_25)

wether_amplicon_MAE_model_array1 <- as.data.frame(Metrics::mae(actual=wether_data_amp_25$YPMP, predicted=wether_data_amp_25$Array_Model_AndroAge))
wether_amplicon_MAE_model_array1
# Wether MAE = 0.608 years
```



### AMPLICON

# Applying array data to amplicon clocks
#### 5 % error clock
```{r}
sheep_array_MKLN1 <- sheep_array_data_ear[,-c(7:15)]
new_colnames <- c("Sample_ID", "Age", "Sex", "BloodSex", "Genotype", "Beta", "YPMP")
colnames(sheep_array_MKLN1) <- new_colnames

sheep_array_MKLN1$Beta <- as.numeric(sheep_array_MKLN1$Beta)

ram_array_ear_MKLN1_data <- sheep_array_MKLN1 %>% filter(Sex=="Ram")
wether_array_ear_MKLN1_data <- sheep_array_MKLN1 %>% filter(Sex=="Wether")
ewe_array_ear_MKLN1_data <- sheep_array_MKLN1 %>% filter(Sex=="Ewe")
  
  
sheep_array_MKLN1$Amplicon_Model_AndroAge_5 <- predict(model_5p, newdata = sheep_array_MKLN1)

sheep_array_MAE_model_5p <- as.data.frame(Metrics::mae(actual=sheep_array_MKLN1$YPMP, predicted=sheep_array_MKLN1$Amplicon_Model_AndroAge_5))
sheep_array_MAE_model_5p
# OVERALL MAE = 0.678 years

sheep_array_MKLN1 <- sheep_array_MKLN1 %>% 
  arrange(Age) %>% 
  mutate(index = row_number())


cor(sheep_array_MKLN1$YPMP, sheep_array_MKLN1$Amplicon_Model_AndroAge_5, method="pearson")


compare_5_array <- ggplot(sheep_array_MKLN1, aes(x=YPMP, y=Amplicon_Model_AndroAge_5, colour=Sex)) + 
  geom_point(size=4, alpha=3/4) +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  geom_smooth(method="lm", se=F, colour="#1F6683") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Train: ±5% error amplicon, Test: array",
              subtitle = "MAE=0.678, cor=0.870") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
compare_5_array
```


```{r}
ram_array_ear_MKLN1_data$Amplicon_Model_AndroAge_5 <- predict(model_5p, newdata = ram_array_ear_MKLN1_data)

ram_array_MAE_model_5p <- as.data.frame(Metrics::mae(actual=ram_array_ear_MKLN1_data$YPMP, predicted=ram_array_ear_MKLN1_data$Amplicon_Model_AndroAge_5))
ram_array_MAE_model_5p
# Ram MAE = 0.810 years

ewe_array_ear_MKLN1_data$Amplicon_Model_AndroAge_5 <- predict(model_5p, newdata = ewe_array_ear_MKLN1_data)

ewe_array_MAE_model_5p <- as.data.frame(Metrics::mae(actual=ewe_array_ear_MKLN1_data$YPMP, predicted=ewe_array_ear_MKLN1_data$Amplicon_Model_AndroAge_5))
ewe_array_MAE_model_5p
# Ewe MAE = 0.595 years

wether_array_ear_MKLN1_data$Amplicon_Model_AndroAge_5 <- predict(model_5p, newdata = wether_array_ear_MKLN1_data)

wether_array_MAE_model_5p <- as.data.frame(Metrics::mae(actual=wether_array_ear_MKLN1_data$YPMP, predicted=wether_array_ear_MKLN1_data$Amplicon_Model_AndroAge_5))
wether_array_MAE_model_5p
# Wether MAE = 0.728 years
```


#### 2.5 % error clock
```{r}
sheep_array_MKLN1$Amplicon_Model_AndroAge_25 <- predict(model_25p, newdata = sheep_array_MKLN1)

sheep_array_MAE_model_25p <- as.data.frame(Metrics::mae(actual=sheep_array_MKLN1$YPMP, predicted=sheep_array_MKLN1$Amplicon_Model_AndroAge_25))
sheep_array_MAE_model_25p
# OVERALL MAE = 0.676 years

sheep_array_MKLN1 <- sheep_array_MKLN1 %>% 
  arrange(Age) %>% 
  mutate(index = row_number())

cor(sheep_array_MKLN1$YPMP, sheep_array_MKLN1$Amplicon_Model_AndroAge_25, method="pearson")

compare_25_array <- ggplot(sheep_array_MKLN1, aes(x=YPMP, y=Amplicon_Model_AndroAge_25, colour=Sex)) + 
  geom_point(size=4, alpha=3/4) +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  geom_smooth(method="lm", se=F, colour="#1F6683") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Train: ±2.5% error amplicon, Test: array",
              subtitle = "MAE=0.676, cor=0.870") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
compare_25_array


```


```{r}
ram_array_ear_MKLN1_data$Amplicon_Model_AndroAge_25 <- predict(model_25p, newdata = ram_array_ear_MKLN1_data)

ram_array_MAE_model_25p <- as.data.frame(Metrics::mae(actual=ram_array_ear_MKLN1_data$YPMP, predicted=ram_array_ear_MKLN1_data$Amplicon_Model_AndroAge_25))
ram_array_MAE_model_25p
# Ram MAE = 0.787 years

ewe_array_ear_MKLN1_data$Amplicon_Model_AndroAge_25 <- predict(model_25p, newdata = ewe_array_ear_MKLN1_data)

ewe_array_MAE_model_25p <- as.data.frame(Metrics::mae(actual=ewe_array_ear_MKLN1_data$YPMP, predicted=ewe_array_ear_MKLN1_data$Amplicon_Model_AndroAge_25))
ewe_array_MAE_model_25p
# Ewe MAE = 0.602 years

wether_array_ear_MKLN1_data$Amplicon_Model_AndroAge_25 <- predict(model_25p, newdata = wether_array_ear_MKLN1_data)

wether_array_MAE_model_25p <- as.data.frame(Metrics::mae(actual=wether_array_ear_MKLN1_data$YPMP, predicted=wether_array_ear_MKLN1_data$Amplicon_Model_AndroAge_25))
wether_array_MAE_model_25p
# Wether MAE = 0.726 years
```


```{r}
compare_array_amplicon <- cowplot::plot_grid(compare_array_5, compare_array_25, compare_5_array, compare_25_array, labels=c("A", "B", "C", "D"))
compare_array_amplicon

legend_b <- cowplot::get_legend(
  compare_5_array + 
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")
)

compare_array_amplicon <- cowplot::plot_grid(compare_array_amplicon,legend_b, ncol=1, rel_heights=c(1,0.1))
compare_array_amplicon

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/array_amplicon_modelfittingcompare_ALL.png",compare_array_amplicon, width=10, height = 10, dpi=1000)

```

# Compare methylation on the array vs amplicon
```{r}
compare <- read.csv("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Bioinformatics/Data/SheepMethylation_AmpliconvsArray.csv")

compare_5 <- compare %>% filter(CG_Calls>=497)
compare_5$Sex <- ifelse(substr(compare_5$Sample, 1, 3) %in% c("Ram", "RAM"), "Ram",
                 ifelse(substr(compare_5$Sample, 1, 3) == "BHC", "Ewe",
                        ifelse(substr(compare_5$Sample, 1, 3) %in% c("Wet", "WET"), "Wether", NA)))


comparison_plot <-ggplot(compare_5, aes(y=CG_Methylation, x=(Array*100), colour=Sex)) + 
  geom_point(size=4, alpha=2/3) +
  theme_light() +
  geom_abline(linetype="dashed") +
  geom_smooth(method="lm", se=F, colour="Black") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  ylab("Amplicon methylation (%)") +
  xlab("Array methylation (%)") +
  theme(legend.position="none")
comparison_plot

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/compareMethylation_AmpliconArray.png",comparison_plot,width=8,height=8)


compare_5_complete <- compare_5[complete.cases(compare_5$Array),]

cor(compare_5_complete$Array, compare_5_complete$CG_Methylation, method="pearson")
# 0.912

compare_5_complete$difference <- compare_5_complete$CG_Methylation - (compare_5_complete$Array*100)
mean(as.numeric(compare_5_complete$difference), na.rm = TRUE)
# 4.23% 
```

```{r}
MKLN1_sheep_amp_5 <- ggplot(sheep_data_amp_5, aes(x=Age, y=CG_Methylation, colour=Castration_Sex)) +
  geom_jitter(width=0.1, size=4, alpha=2/3) +
  geom_smooth(method="lm") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(colour="Sex", y="Methylation (%)", x="Age (Years)") +
  theme(legend.position="none") +
  ylim(30,90)

MKLN1_amplicon_corr <- cowplot::plot_grid(MKLN1_sheep_amp_5, comparison_plot, nrow=1, labels=c("A", "B"))
MKLN1_amplicon_corr

legend_b <- get_legend(
  MKLN1_sheep_amp_5 + 
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")
)

MKLN1_amplicon_corr_legend <- cowplot::plot_grid(MKLN1_amplicon_corr, legend_b, ncol=1, rel_heights=c(1,0.08))
MKLN1_amplicon_corr_legend

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/MKLN1_array_amplicon_MethPlatform_2.png",MKLN1_amplicon_corr_legend, width=10, height = 6, dpi=1000)

```
