---
title: "Sheep Androgen Clock Construction - BBA-seq approach"
author: "Victoria Sugrue"
date: 'Last updated: 15 May 2023'
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 6
  pdf_document:
    toc: yes
    toc_depth: 5
    df_print: kable
    latex_engine: xelatex
subtitle: ' '
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r}
library(ggplot2)
library(caret)
library(dplyr)
library(Manu)
library(knitr)
library(sjPlot)
library(lmtest)
library(Amelia)
library(mlbench)
library(ggfortify)
```

# Data preparation
```{r}
sheep_data <- read.csv("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Bioinformatics/Data/SheepSampleSheet.csv")

# Checking for missingness
#pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/sheepdata_missingness.pdf") 
missmap(sheep_data, col=c("blue", "red"), legend=T)
#dev.off()

# Setting AndroAge with predicted time of puberty onset = age minus 0.5 years (i.e estimate to reach puberty at 6 months)
sheep_data$YPMP <- sheep_data$Age-0.5
sheep_data$YPMP[sheep_data$Castration_Sex=="Wether"] <- 0
sheep_data$YPMP[sheep_data$Castration_Sex=="Ewe"] <- 0

# Creating a new column for methylation in the form of a beta value
sheep_data$Beta <- sheep_data$CG_Methylation/100

# Subset by sex/castration status
ram_data <- sheep_data %>% filter(Castration_Sex=="Ram")
wether_data <- sheep_data %>% filter(Castration_Sex=="Wether")
ewe_data <- sheep_data %>% filter(Castration_Sex=="Ewe")

```
# A birds-eye view of the data 

```{r}
# Age distribution
age_dist <- ggplot(sheep_data, aes(x = Age)) +
  geom_histogram(aes(fill = Castration_Sex), position = "dodge", alpha = 1, bins = 20) +
  scale_fill_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Distribution of age by sex (Sheep ear)", fill = "Sex", x="Age (Years)") +
  ylab("Frequency")
age_dist

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/age_distribution_clockSheep.png",age_dist, width=10, height = 8, dpi=1000)
```

```{r}
# Subset based on error allowance
sheep_data_5 <- sheep_data %>% filter(CG_Calls>=497)
sheep_data_25 <- sheep_data %>% filter(CG_Calls>=1990)
sheep_data_2 <- sheep_data %>% filter(CG_Calls>=3110)
sheep_data_15 <- sheep_data %>% filter(CG_Calls>=5529)
sheep_data_1 <- sheep_data %>% filter(CG_Calls>=12440)

# Age distribution with +/- 5% error in 75% methylated samples
age_dist_5 <- ggplot(sheep_data_5, aes(x = Age)) +
  geom_histogram(aes(fill = Castration_Sex), position = "dodge", alpha = 1, bins = 20) +
  scale_fill_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Distribution of age by sex (Sheep ear), >497 CG calls (±5% error at 75% methyated)", fill = "Sex", x="Age (Years)") +
  ylab("Frequency")
age_dist_5

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/age_distribution_clockSheep_5.png",age_dist_5, width=10, height = 8, dpi=1000)
```

```{r}
# Age distribution with +/- 2.5% error in 75% methylated samples
age_dist_25 <- ggplot(sheep_data_25, aes(x = Age)) +
  geom_histogram(aes(fill = Castration_Sex), position = "dodge", alpha = 1, bins = 20) +
  scale_fill_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Distribution of age by sex (Sheep ear), >1990 CG calls (±2.5% error at 75% methyated)", fill = "Sex", x="Age (Years)") +
  ylab("Frequency")
age_dist_25

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/age_distribution_clockSheep_25.png",age_dist_25, width=10, height = 8, dpi=1000)
```

```{r}
# Age distribution with +/- 2% error in 75% methylated samples
age_dist_2 <- ggplot(sheep_data_2, aes(x = Age)) +
  geom_histogram(aes(fill = Castration_Sex), position = "dodge", alpha = 1, bins = 20) +
  scale_fill_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Distribution of age by sex (Sheep ear), >3110 CG calls (±2% error at 75% methyated)", fill = "Sex", x="Age (Years)") +
  ylab("Frequency")
age_dist_2

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/age_distribution_clockSheep_2.png",age_dist_2, width=10, height = 8, dpi=1000)
```

```{r}
# Age distribution with +/- 1.5% error in 75% methylated samples
age_dist_15 <- ggplot(sheep_data_15, aes(x = Age)) +
  geom_histogram(aes(fill = Castration_Sex), position = "dodge", alpha = 1, bins = 20) +
  scale_fill_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Distribution of age by sex (Sheep ear), >5529 CG calls (±1.5% error at 75% methyated)", fill = "Sex", x="Age (Years)") +
  ylab("Frequency")
age_dist_15

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/age_distribution_clockSheep_15.png",age_dist_15, width=10, height = 8, dpi=1000)
```

```{r}
# Age distribution with +/- 1% error in 75% methylated samples
age_dist_1 <- ggplot(sheep_data_1, aes(x = Age)) +
  geom_histogram(aes(fill = Castration_Sex), position = "dodge", alpha = 1, bins = 20) +
  scale_fill_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Distribution of age by sex (Sheep ear), >12440 CG calls (±1% error at 75% methyated)", fill = "Sex", x="Age (Years)") +
  ylab("Frequency")
age_dist_1

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/age_distribution_clockSheep_1.png",age_dist_1, width=10, height = 8, dpi=1000)
```

# Plotting methylation data

```{r}
# +/- 5% error in 75% methylated samples
MKLN1_sheep_5 <- ggplot(sheep_data_5, aes(x=Age, y=CG_Methylation, colour=Castration_Sex)) +
  geom_jitter(width=0.1, size=4, alpha=2/3) +
  geom_smooth(method="lm") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Methylation at cg21524116, >497 CG calls (±5% error at 75% methyated)", colour="Sex", y="Methylation (%)", x="Age (Years)")
MKLN1_sheep_5

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/MKLN1_meth_sheepear_5.png",MKLN1_sheep_5, width=10, height = 8, dpi=1000)
```


```{r}
# +/- 2.5% error in 75% methylated samples
MKLN1_sheep_25 <- ggplot(sheep_data_25, aes(x=Age, y=CG_Methylation, colour=Castration_Sex)) +
  geom_jitter(width=0.1, size=4, alpha=2/3) +
  geom_smooth(method="lm") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Methylation at cg21524116, >1990 CG calls (±2.5% error at 75% methyated)", colour="Sex", y="Methylation (%)", x="Age (Years)")
MKLN1_sheep_25

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/MKLN1_meth_sheepear_25.png",MKLN1_sheep_25, width=10, height = 8, dpi=1000)
```

```{r}
# +/- 2% error in 75% methylated samples
MKLN1_sheep_2 <- ggplot(sheep_data_2, aes(x=Age, y=CG_Methylation, colour=Castration_Sex)) +
  geom_jitter(width=0.1, size=4, alpha=2/3) +
  geom_smooth(method="lm") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Methylation at cg21524116, >3110 CG calls (±2% error at 75% methyated)", colour="Sex", y="Methylation (%)", x="Age (Years)")
MKLN1_sheep_2

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/MKLN1_meth_sheepear_2.png",MKLN1_sheep_2, width=10, height = 8, dpi=1000)
```

```{r}
# +/- 1.5% error in 75% methylated samples
MKLN1_sheep_15 <- ggplot(sheep_data_15, aes(x=Age, y=CG_Methylation, colour=Castration_Sex)) +
  geom_jitter(width=0.1, size=4, alpha=2/3) +
  geom_smooth(method="lm") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Methylation at cg21524116, >5529 CG calls (±1.5% error at 75% methyated)", colour="Sex", y="Methylation (%)", x="Age (Years)")
MKLN1_sheep_15

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/MKLN1_meth_sheepear_15.png",MKLN1_sheep_15, width=10, height = 8, dpi=1000)
```

```{r}
MKLN1_sheep_1 <- ggplot(sheep_data_1, aes(x=Age, y=CG_Methylation, colour=Castration_Sex)) +
  geom_jitter(width=0.1, size=4, alpha=2/3) +
  geom_smooth(method="lm") +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme_light() +
  labs(title = "Methylation at cg21524116, >12440 CG calls (±1% error at 75% methyated)", colour="Sex", y="Methylation (%)", x="Age (Years)")
MKLN1_sheep_1

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/MKLN1_meth_sheepear_1.png",MKLN1_sheep_1, width=10, height = 8, dpi=1000)
```


# Fitting the Androgen Clock to rams

```{r}
ram_data_5 <- ram_data %>% filter(CG_Calls>=497)
ram_data_25 <- ram_data %>% filter(CG_Calls>=1990)
ram_data_15 <- ram_data %>% filter(CG_Calls>=5529)
ram_data_2 <- ram_data %>% filter(CG_Calls>=3110)
ram_data_1 <- ram_data %>% filter(CG_Calls>=12440)
```


```{r}
# Run model, 5% error
model_5p <- lm(YPMP ~ Beta, data=ram_data_5)
summary(model_5p)
```

# Diagnostic plots

```{r}
pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/Model5p_ggplot_diagnostics_white.pdf", width=10, height=8)

autoplot(model_5p, which=c(1,2,3,5), size=3, alpha=4/5, colour="#313657", smooth.colour="#DD3C51", smooth.linetype="dashed", label.size = 3) + theme_light() + theme(plot.title = element_text(color = "white"))

dev.off()
```


```{r}
fitted_5 <- as.data.frame(model_5p$fitted.values)
fitted_5["YPMP"] <- ram_data_5$YPMP
fitted_5$YPMP <- as.numeric(fitted_5$YPMP)
fitted_5$Fit_AndroAge <- as.numeric(fitted_5$`model_5p$fitted.values`)
fitted_5$Age <- ram_data_5$Age
```

# Calculating error
## All

```{r}
MAE_model_5p <- as.data.frame(Metrics::mae(actual=fitted_5$YPMP, predicted=fitted_5$`model_5p$fitted.values`))
MAE_model_5p # 0.569
```

# Plot 

```{r}
cor(fitted_5$YPMP, fitted_5$Fit_AndroAge, method="pearson")

ramfit5 <- ggplot(fitted_5, aes(x=YPMP, y=Fit_AndroAge)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Ram androgen clock fit (model_5p), trained on all",
              subtitle = "MAE=0.569, cor=0.912") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))

ramfit5
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/5errormodel_ramFit.png",ramfit5, width=8, height = 7, dpi=1000)

```


# Testing the model fit further: 

# Test density/normality of residuals - looking for normally distributed. 
```{r}
d<-density(model_5p[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

```{r}
ggplot(data = ram_data_5, aes(x = model_5p$residuals)) +
    geom_histogram(fill = 'grey40', color = 'black', bins=20) +
    labs(title = 'Histogram of Residuals, model 1 (5% error)', x = 'Residuals', y = 'Frequency')
```
# Breusch-Pagan test of variance (homo/heteroscedasticity)

```{r}
plot(model_5p$residuals)
bptest(model_5p)
```

# 0.06
# p value >0.05 suggests that we have homoscedasticity. good! 

# Shapiro-Wilk test of normality

```{r}
shapiro.test(model_5p$residuals)
```

# 0.90
# p value >0.05 suggests data is normally distributed. good!

# Leave-one-out cross validation

```{r}
ram_data_5$Beta <- as.numeric(ram_data_5$Beta)
ram_data_5$YPMP <- as.numeric(ram_data_5$YPMP)
ctrl <- trainControl(method="LOOCV")
model_5p_LOO <- train(YPMP ~ Beta, data=ram_data_5, method="lm", trControl=ctrl)
summary(model_5p_LOO)
```

```{r}
print(model_5p_LOO)
```

```{r}
fittedLOO_5 <- as.data.frame(model_5p_LOO$pred)
fittedLOO_5["YPMP"] <- ram_data_5$YPMP

LOORamAndroClockMAE_5 <- as.data.frame(Metrics::mae(actual=fittedLOO_5$YPMP, predicted=fittedLOO_5$pred))
LOORamAndroClockMAE_5 # 0.579
```

```{r}
cor(fittedLOO_5$obs, fittedLOO_5$pred, method="pearson")

ramLOOfit5 <- ggplot(fittedLOO_5, aes(x=obs, y=pred)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Ram androgen clock fit (model_5p), LOOCV",
              subtitle = "MAE=0.579, cor=0.909") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
ramLOOfit5
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/5errormodel_LOO_ramFit.png",ramLOOfit5, width=8, height = 7, dpi=1000)

```

# Now, to include ewes and wethers.

```{r}
ewe_data_5 <- ewe_data %>% filter(CG_Calls>=497)
ewe_data_5$Model_AndroAge <- predict(model_5p, newdata = ewe_data_5)

ewe_MAE_model_5p <- as.data.frame(Metrics::mae(actual=ewe_data_5$YPMP, predicted=ewe_data_5$Model_AndroAge))
ewe_MAE_model_5p
# Ewe MAE = 0.38 years
```


```{r}
wether_data_5 <- wether_data %>% filter(CG_Calls>=497)
wether_data_5$Model_AndroAge <- predict(model_5p, newdata = wether_data_5)

wether_MAE_model_5p <- as.data.frame(Metrics::mae(actual=wether_data_5$YPMP, predicted=wether_data_5$Model_AndroAge))
wether_MAE_model_5p
# Wether MAE = 0.33 years

ewewether_5 <- rbind(ewe_data_5, wether_data_5)

ewewether_5_violin <- ggplot(ewewether_5, aes(x=Castration_Sex, y=Model_AndroAge, fill=Castration_Sex)) +
  geom_violin() + 
  geom_point(size=4, alpha=3/4, shape=1) +
  theme_light() +
  geom_hline(yintercept = 0, linetype="dashed") +
  ylab("Model AndroAge (Years)") +
  xlab("Sex") +
  scale_fill_manual(values = get_pal("Takahe")[c(1,4)]) +
  theme(legend.position = "none") +
   labs(title = "Ewe and wether androgen clock fit (model_5p), trained on all rams",
              subtitle = "Ewe MAE=0.389, Wether MAE=0.331") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
  
ewewether_5_violin

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/5errormodel_ewewetherFit_violin.png",ewewether_5_violin, width=11, height = 6, dpi=1000)

```

# Running model, 2.5% error

```{r}
# Run model, 2.5% error
model_25p <- lm(YPMP ~ Beta, data=ram_data_25)
summary(model_25p)
```

# Diagnostic plots

```{r}
pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/Model25p_ggplot_diagnostics_white.pdf", width=10, height=8)

autoplot(model_25p, which=c(1,2,3,5), size=3, alpha=4/5, colour="#313657", smooth.colour="#DD3C51", smooth.linetype="dashed", label.size = 3) + theme_light() + theme(plot.title = element_text(color = "white"))

dev.off()
```


```{r}
fitted_25 <- as.data.frame(model_25p$fitted.values)
fitted_25["YPMP"] <- ram_data_25$YPMP
fitted_25$YPMP <- as.numeric(fitted_25$YPMP)
fitted_25$Fit_AndroAge <- as.numeric(fitted_25$`model_25p$fitted.values`)
```

# Calculating error
## All

```{r}
MAE_model_25p <- as.data.frame(Metrics::mae(actual=fitted_25$YPMP, predicted=fitted_25$`model_25p$fitted.values`))
MAE_model_25p # 0.535
```
# Plot 

```{r}
cor(fitted_25$YPMP, fitted_25$Fit_AndroAge, method="pearson")

ramfit25 <- ggplot(fitted_25, aes(x=YPMP, y=Fit_AndroAge)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
    labs(title = "Ram androgen clock fit (model_25p), trained on all",
              subtitle = "MAE=0.535, cor=0.906") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
ramfit25
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/25errormodel_ramFit.png",ramfit25, width=8, height = 7, dpi=1000)
```

# Testing the model fit further: 

# Test density/normality of residuals - looking for normally distributed. 
```{r}
d<-density(model_25p[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

```{r}
ggplot(data = ram_data_25, aes(x = model_25p$residuals)) +
    geom_histogram(fill = 'grey40', color = 'black', bins=20) +
    labs(title = 'Histogram of Residuals, model 1 (2.5% error)', x = 'Residuals', y = 'Frequency')
```
# Breusch-Pagan test of variance (homo/heteroscedasticity)

```{r}
plot(model_25p$residuals)
bptest(model_25p)
```

# 0.001
# p value <0.05 suggests that we have heteroscedasticity. not good! 

# Shapiro-Wilk test of normality

```{r}
shapiro.test(model_25p$residuals)
```

# p value >0.05 suggests data is normally distributed. good!

# Leave-one-out cross validation

```{r}
ram_data_25$Beta <- as.numeric(ram_data_25$Beta)
ram_data_25$YPMP <- as.numeric(ram_data_25$YPMP)
ctrl <- trainControl(method="LOOCV")
model_25p_LOO <- train(YPMP ~ Beta, data=ram_data_25, method="lm", trControl=ctrl)
summary(model_25p_LOO)
```

```{r}
print(model_25p_LOO)
```

```{r}
fittedLOO_25 <- as.data.frame(model_25p_LOO$pred)
fittedLOO_25["YPMP"] <- ram_data_25$YPMP

LOORamAndroClockMAE_25 <- as.data.frame(Metrics::mae(actual=fittedLOO_25$YPMP, predicted=fittedLOO_25$pred))
LOORamAndroClockMAE_25 # 0.568
```

```{r}
cor(fittedLOO_25$obs, fittedLOO_25$pred, method="pearson")

ramLOOfit25 <- ggplot(fittedLOO_25, aes(x=obs, y=pred)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Ram androgen clock fit (model_25p), LOOCV",
              subtitle = "MAE=0.568, cor=0.892") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
ramLOOfit25
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/25errormodel_LOO_ramFit.png",ramLOOfit25, width=8, height = 7, dpi=1000)

```

# Now, to include ewes and wethers.

```{r}
ewe_data_25 <- ewe_data %>% filter(CG_Calls>=497)
ewe_data_25$Model_AndroAge <- predict(model_25p, newdata = ewe_data_25)

ewe_MAE_model_25p <- as.data.frame(Metrics::mae(actual=ewe_data_25$YPMP, predicted=ewe_data_25$Model_AndroAge))
ewe_MAE_model_25p
# Ewe MAE = 0.35 years
```


```{r}
wether_data_25 <- wether_data %>% filter(CG_Calls>=497)
wether_data_25$Model_AndroAge <- predict(model_25p, newdata = wether_data_25)

wether_MAE_model_25p <- as.data.frame(Metrics::mae(actual=wether_data_25$YPMP, predicted=wether_data_25$Model_AndroAge))
wether_MAE_model_25p
# Wether MAE = 0.31 years

ewewether_25 <- rbind(ewe_data_25, wether_data_25)

ewewether_25_violin <- ggplot(ewewether_25, aes(x=Castration_Sex, y=Model_AndroAge, fill=Castration_Sex)) +
  geom_violin() + 
  geom_point(size=4, alpha=3/4, shape=1) +
  theme_light() +
  geom_hline(yintercept = 0, linetype="dashed") +
  ylab("Model AndroAge (Years)") +
  xlab("Sex") +
  scale_fill_manual(values = get_pal("Takahe")[c(1,4)]) +
  theme(legend.position = "none") +
   labs(title = "Ewe and wether androgen clock fit (model_25p), trained on all rams",
              subtitle = "Ewe MAE=0.351, Wether MAE=0.317") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
  
ewewether_25_violin

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/25errormodel_ewewetherFit_violin.png",ewewether_25_violin, width=11, height = 6, dpi=1000)

```

# Running model, 2% error

```{r}
# Run model, 2% error
model_2p <- lm(YPMP ~ Beta, data=ram_data_2)
summary(model_2p)
```

# Diagnostic plots
```{r}
pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/Model2p_ggplot_diagnostics_white.pdf", width=10, height=8)

autoplot(model_2p, which=c(1,2,3,5), size=3, alpha=4/5, colour="#313657", smooth.colour="#DD3C51", smooth.linetype="dashed", label.size = 3) + theme_light() + theme(plot.title = element_text(color = "white"))

dev.off()

```


```{r}
fitted_2 <- as.data.frame(model_2p$fitted.values)
fitted_2["YPMP"] <- ram_data_2$YPMP
fitted_2$YPMP <- as.numeric(fitted_2$YPMP)
fitted_2$Fit_AndroAge <- as.numeric(fitted_2$`model_2p$fitted.values`)
```

# Calculating error
## All

```{r}
MAE_model_2p <- as.data.frame(Metrics::mae(actual=fitted_2$YPMP, predicted=fitted_2$`model_2p$fitted.values`))
MAE_model_2p # 0.490
```

# Plot 

```{r}
cor(fitted_2$YPMP, fitted_2$Fit_AndroAge, method="pearson")

ramfit2 <- ggplot(fitted_2, aes(x=YPMP, y=Fit_AndroAge)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("AndroAge predicted by model (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Ram androgen clock fit (model_2p), trained on all",
              subtitle = "MAE=0.490, cor=0.908") +
    theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
ramfit2
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/2errormodel_ramFit.png",ramfit2, width=8, height = 7, dpi=1000)

```

# Testing the model fit further: 

# Test density/normality of residuals - looking for normally distributed. 
```{r}
d<-density(model_2p[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

```{r}
ggplot(data = ram_data_2, aes(x = model_2p$residuals)) +
    geom_histogram(fill = 'grey40', color = 'black', bins=20) +
    labs(title = 'Histogram of Residuals, model 1 (1.5% error)', x = 'Residuals', y = 'Frequency')
```
# Breusch-Pagan test of variance (homo/heteroscedasticity)

```{r}
plot(model_2p$residuals)
bptest(model_2p)
```

# p value <0.05 suggests that we have heteroscedasticity. not good! 

# Shapiro-Wilk test of normality

```{r}
shapiro.test(model_2p$residuals)
```

# p value >0.05 suggests data is normally distributed. good!

# Leave-one-out cross validation

```{r}
ram_data_2$Beta <- as.numeric(ram_data_2$Beta)
ram_data_2$YPMP <- as.numeric(ram_data_2$YPMP)
ctrl <- trainControl(method="LOOCV")
model_2p_LOO <- train(YPMP ~ Beta, data=ram_data_2, method="lm", trControl=ctrl)
summary(model_2p_LOO)
```

```{r}
print(model_2p_LOO)
```

```{r}
fittedLOO_2 <- as.data.frame(model_2p_LOO$pred)
fittedLOO_2["YPMP"] <- ram_data_2$YPMP

LOORamAndroClockMAE_2 <- as.data.frame(Metrics::mae(actual=fittedLOO_2$YPMP, predicted=fittedLOO_2$pred))
LOORamAndroClockMAE_2 # 0.535
```

```{r}
cor(fittedLOO_2$obs, fittedLOO_2$pred, method="pearson")

ramLOOfit2 <- ggplot(fittedLOO_2, aes(x=obs, y=pred)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Ram androgen clock fit (model_2p), LOOCV",
              subtitle = "MAE=0.535, cor=0.888") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
ramLOOfit2
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/2errormodel_LOO_ramFit.png",ramLOOfit2, width=8, height = 7, dpi=1000)

```

# Now, to include ewes and wethers.

```{r}
ewe_data_2 <- ewe_data %>% filter(CG_Calls>=497)
ewe_data_2$Model_AndroAge <- predict(model_2p, newdata = ewe_data_2)

ewe_MAE_model_2p <- as.data.frame(Metrics::mae(actual=ewe_data_2$YPMP, predicted=ewe_data_2$Model_AndroAge))
ewe_MAE_model_2p
# Ewe MAE = 0.54 years
```


```{r}
wether_data_2 <- wether_data %>% filter(CG_Calls>=497)
wether_data_2$Model_AndroAge <- predict(model_2p, newdata = wether_data_5)

wether_MAE_model_2p <- as.data.frame(Metrics::mae(actual=wether_data_2$YPMP, predicted=wether_data_2$Model_AndroAge))
wether_MAE_model_2p
# Wether MAE = 0.40 years

ewewether_2 <- rbind(ewe_data_2, wether_data_2)

ewewether_2_violin <- ggplot(ewewether_2, aes(x=Castration_Sex, y=Model_AndroAge, fill=Castration_Sex)) +
  geom_violin() + 
  geom_point(size=4, alpha=3/4, shape=1) +
  theme_light() +
  geom_hline(yintercept = 0, linetype="dashed") +
  ylab("Model AndroAge (Years)") +
  xlab("Sex") +
  scale_fill_manual(values = get_pal("Takahe")[c(1,4)]) +
  theme(legend.position = "none") +
   labs(title = "Ewe and wether androgen clock fit (model_2p), trained on all rams",
              subtitle = "Ewe MAE=0.547, Wether MAE=0.408") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
  
ewewether_2_violin

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/2errormodel_ewewetherFit_violin.png",ewewether_2_violin, width=11, height = 6, dpi=1000)

```


# Running model, 1.5% error


```{r}
# Run model, 1.5% error
model_15p <- lm(YPMP ~ Beta, data=ram_data_15)
summary(model_15p)
```

# Diagnostic plots
```{r}
pdf("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/Model15p_ggplot_diagnostics_white.pdf", width=10, height=8)

autoplot(model_15p, which=c(1,2,3,5), size=3, alpha=4/5, colour="#313657", smooth.colour="#DD3C51", smooth.linetype="dashed", label.size = 3) + theme_light() + theme(plot.title = element_text(color = "white"))

dev.off()

```


```{r}
fitted_15 <- as.data.frame(model_15p$fitted.values)
fitted_15["YPMP"] <- ram_data_15$YPMP
fitted_15$YPMP <- as.numeric(fitted_15$YPMP)
fitted_15$Fit_AndroAge <- as.numeric(fitted_15$`model_15p$fitted.values`)
```

# Calculating error
## All

```{r}
MAE_model_15p <- as.data.frame(Metrics::mae(actual=fitted_15$YPMP, predicted=fitted_15$`model_15p$fitted.values`))
MAE_model_15p # 0.499
```

# Plot 

```{r}
cor(fitted_15$YPMP, fitted_15$Fit_AndroAge, method="pearson")

ramfit15 <- ggplot(fitted_15, aes(x=YPMP, y=Fit_AndroAge)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("AndroAge predicted by model (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
    labs(title = "Ram androgen clock fit (model_15p), trained on all",
              subtitle = "MAE=0.499, cor=0.924") +
    theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
ramfit15
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/15errormodel_ramFit.png",ramfit15, width=8, height = 7, dpi=1000)

```

# Testing the model fit further: 

# Test density/normality of residuals - looking for normally distributed. 
```{r}
d<-density(model_15p[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

```{r}
ggplot(data = ram_data_15, aes(x = model_15p$residuals)) +
    geom_histogram(fill = 'grey40', color = 'black', bins=20) +
    labs(title = 'Histogram of Residuals, model 1 (1.5% error)', x = 'Residuals', y = 'Frequency')
```
# Breusch-Pagan test of variance (homo/heteroscedasticity)

```{r}
plot(model_15p$residuals)
bptest(model_15p)
```

# p value >0.05 suggests that we have homoscedasticity. good! 

# Shapiro-Wilk test of normality

```{r}
shapiro.test(model_15p$residuals)
```

# p value >0.05 suggests data is normally distributed. good!

# Leave-one-out cross validation

```{r}
ram_data_15$Beta <- as.numeric(ram_data_15$Beta)
ram_data_15$YPMP <- as.numeric(ram_data_15$YPMP)
ctrl <- trainControl(method="LOOCV")
model_15p_LOO <- train(YPMP ~ Beta, data=ram_data_15, method="lm", trControl=ctrl)
summary(model_15p_LOO)
```

```{r}
print(model_15p_LOO)
```

```{r}
fittedLOO_15 <- as.data.frame(model_15p_LOO$pred)
fittedLOO_15["YPMP"] <- ram_data_15$YPMP

LOORamAndroClockMAE_15 <- as.data.frame(Metrics::mae(actual=fittedLOO_15$YPMP, predicted=fittedLOO_15$pred))
LOORamAndroClockMAE_15 # 0.599
```

```{r}
cor(fittedLOO_15$obs, fittedLOO_15$pred, method="pearson")

ramLOOfit15 <- ggplot(fittedLOO_15, aes(x=obs, y=pred)) +
  geom_point(size=4, alpha=3/4, colour="#313657") + 
  geom_smooth(method="lm", se=F, colour="#DD3C51") +
  geom_abline(linetype="dashed") +
  theme_light() +
  ylab("Model AndroAge (Years)") +
  xlab("Years Post Male Puberty (YPMP)") +
  labs(title = "Ram androgen clock fit (model_15p), LOOCV",
              subtitle = "MAE=599, cor=0.894") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
ramLOOfit15
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/15errormodel_LOO_ramFit.png",ramLOOfit15, width=8, height = 7, dpi=1000)

```

# Now, to include ewes and wethers.

```{r}
ewe_data_15 <- ewe_data %>% filter(CG_Calls>=497)
ewe_data_15$Model_AndroAge <- predict(model_15p, newdata = ewe_data_15)

ewe_MAE_model_15p <- as.data.frame(Metrics::mae(actual=ewe_data_15$YPMP, predicted=ewe_data_15$Model_AndroAge))
ewe_MAE_model_15p
# Ewe MAE = 0.49 years
```


```{r}
wether_data_15 <- wether_data %>% filter(CG_Calls>=497)
wether_data_15$Model_AndroAge <- predict(model_15p, newdata = wether_data_15)

wether_MAE_model_15p <- as.data.frame(Metrics::mae(actual=wether_data_15$YPMP, predicted=wether_data_15$Model_AndroAge))
wether_MAE_model_15p
# Wether MAE = 0.38 years

ewewether_15 <- rbind(ewe_data_15, wether_data_15)

ewewether_15_violin <- ggplot(ewewether_15, aes(x=Castration_Sex, y=Model_AndroAge, fill=Castration_Sex)) +
  geom_violin() + 
  geom_point(size=4, alpha=3/4, shape=1) +
  theme_light() +
  geom_hline(yintercept = 0, linetype="dashed") +
  ylab("Model AndroAge (Years)") +
  xlab("Sex") +
  scale_fill_manual(values = get_pal("Takahe")[c(1,4)]) +
  theme(legend.position = "none") +
   labs(title = "Ewe and wether androgen clock fit (model_15p), trained on all rams",
              subtitle = "Ewe MAE=0.497, Wether MAE=0.384") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
  
ewewether_15_violin

ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/15errormodel_ewewetherFit_violin.png",ewewether_15_violin, width=11, height = 6, dpi=1000)

```

```{r}
ram_data_5$Model_AndroAge <- predict(model_5p, newdata = ram_data_5)

combined_5 <- rbind(ram_data_5, wether_data_5, ewe_data_5)

combined_5$AndroAgeAccel <- combined_5$Model_AndroAge-combined_5$YPMP

write.csv(combined_5, "/Users/victoriasugrue/Dropbox (Otago University)/combined_5.csv")
d2combined_5 <- read.csv("/Users/victoriasugrue/Dropbox (Otago University)/d2_combined_5.csv")

SheepAndroAgeAccel_Amplicon <- ggplot(d2combined_5, aes(x=Group, y=mean)) + 
  geom_bar(stat = "identity", fill="lightgrey", colour="black") + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.5) + 
  geom_jitter(combined_5, mapping = aes(x=Castration_Sex, y=AndroAgeAccel, colour=Castration_Sex), size=4, alpha=1/2, width=0.05) +
  ylab("AndroAge Acceleration (Years)") + theme_classic() + geom_hline(yintercept=0) + 
  theme_light() +
  scale_colour_manual(values = get_pal("Takahe")[c(1,2,4)]) +
  theme(legend.position="none")
SheepAndroAgeAccel_Amplicon
ggsave("/Users/victoriasugrue/Dropbox (Otago University)/00_ThesisWriting/Figures/SheepAndroAgeAccel_Amplicon.png", SheepAndroAgeAccel_Amplicon, width=5,height=4, dpi=1000)
```
